---
title: "Homework 2"
author: Zhongqi Yue
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```

## Problem 1

Read the Mr. Trashwheel dataset.

```{r}
trashwheel_df = 
  read_xlsx(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "Mr. Trash Wheel",
    range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2018 and 2017.

```{r}
precip_2018 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2018 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2018) %>% 
  relocate(year)

precip_2017 =
  read_excel(
    "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2017 Precipitation",
    skip = 1
    ) %>% 
  janitor::clean_names() %>% 
  drop_na(month) %>% 
  mutate(year = 2017) %>% 
  relocate(year)
```

Now combine annual precipitation.

```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2018,precip_2017) 

left_join(precip_df, month_df, by = "month" )
```

This dataset contains information from the Mr. Trashwheel trash collector in Baltimore, Maryland. As trash enters the inner harbor, the trashwheel collects that trash, and stores it in a dumpster. The dataset contains information on year, month, and trash collected, include some specific kinds of trash. There are a total of `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include month precipitation data.
The median number of sports balls found in a dumpster in 2017 was `r trashwheel_df %>% filter(year == 2017) %>% pull(sports_balls) %>% median()`
The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% pull(total) %>% sum()` inches.

# Problem 2

Read and clean the NYC Transit dataset.

```{r}
NYC_transit_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv")
NYC_transit_df = janitor::clean_names(NYC_transit_df)
```

Choose some columns and not others.

```{r}
select(NYC_transit_df, line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada)
```

Now convert the entry variable from character to a logical variable.

```{r}
NYC_transit_df = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, entry, vending, entrance_type, ada) %>% 
  mutate(entry = recode(entry, "YES"="TRUE", "NO"="FALSE"), entry = as.logical(entry))
```

The original data set contains 32 variables: `r names(read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv"))`, which include character, double and logical variables. First, I clean the name of each variable to make them all in lower case. Then I only retain the line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance, which are the variables the question interested in. After this step, there are only 19 variables left: `r names(NYC_transit_df)`. Next, I convert the "entry" from character variable to logical variable by using "recode" function. The resulting data set has `r nrow(NYC_transit_df)` rows and `r ncol(NYC_transit_df)` columns. These data are tidy now such that columns are variable names all in lower case and single columns contain single variables.

```{r}
distinct_station_df =
  NYC_transit_df%>% 
  distinct(line, station_name, .keep_all = TRUE)


ada_compliant_df = 
  NYC_transit_df %>% 
  filter(ada == "TRUE")

distinct_ada_compliant_df = 
  NYC_transit_df %>% 
  filter(ada == "TRUE") %>% 
  distinct(line, station_name, ada, .keep_all = TRUE) 

count(filter(NYC_transit_df, entry == "TRUE", vending == "NO"))/count(filter(NYC_transit_df, vending == "NO"))
```

* There are `r count(distinct_station_df)` distinct stations.
* `r count(ada_compliant_df)` stations are ADA compliant. Among these, `r count(distinct_ada_compliant_df)`  distinct stations are ADA compliant.
* The proportion of station entrances / exits without vending allow entrance is `r count(filter(NYC_transit_df, entry == "TRUE", vending == "NO"))/count(filter(NYC_transit_df, vending == "NO"))`.

Reformat the data.

```{r}
Reformed_df = 
  NYC_transit_df %>% 
  mutate_at(vars(route1:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_name",
    names_prefix = "route_name",
    values_to = "route_number") %>% 
  drop_na(route_number) 

serve_A_df=
  Reformed_df%>% 
  filter(route_number == "A") %>% 
  distinct(line, station_name) 

serve_A_compliant_df = 
  Reformed_df%>% 
  filter(route_number == "A") %>% 
  distinct(line, station_name, ada) %>%
  filter(ada == "TRUE") 
```

* After reformatting the data set, there are `r count(serve_A_df)` distinct stations serve the A train. Of the stations that serve the A train, only `r count(serve_A_compliant_df)` are ADA compliant.

# Problem 3

Read and clean the pols-month dataset.

```{r}
pols_month_df = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year","month", "day")) %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)
  ) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president =prez_dem, prez_gop)%>% 
  mutate(president = recode(president,"1" = "dem", "0" = "gop")) %>% 
  select(-prez_gop, -prez_dem) %>% 
  select(-day)
```

Read and clean the snp.csv dataset.

```{r}
snp_df = 
  read_csv("./fivethirtyeight_datasets/snp.csv")%>% 
  janitor::clean_names() %>% 
  separate(date, c("month","day", "year")) %>% 
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)
  ) %>% 
  mutate(month = month.name[month]) %>% 
  relocate(year, month) %>% 
  select(-day)
```

Read and clean the unemployment dataset.

```{r}
unemployment_df = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv") 
colnames(unemployment_df)[2:13] = month.name
```

Wide format to long format.

```{r}
unemployment_df_tidy = 
  unemployment_df %>% 
  pivot_longer(
    January : December,
    names_to = "month",
    values_to = "unemployment"
  ) %>% 
  janitor::clean_names()
```

Joining three datasets.

```{r}
pols_snp_df=
  left_join(pols_month_df, snp_df, by =c("year","month"))
pols_snp_unemployment_df =
  left_join(pols_snp_df, unemployment_df_tidy, by =c("year", "month"))
```

* The original pols-month data set contains 9 variables, which are `r names(read_csv("./fivethirtyeight_datasets/pols-month.csv"))`. Among these variables, mon is date and the rest are double variables. After cleaning the data set, mon variable is separated into integer variables year, month and day. Then month number is replaced with the month name, which turns it into a character variable. A president variable is created to present prez_dem and prez_gop variables. Then prez_dem, prez_gop and day variables are removed.The final pols-month data set contains 9 varibales, which are `r names(pols_month_df)`. The final dimension (rows, columns) of this data set is (`r dim(pols_month_df)`). 
* The original snp data set contains two variables, which are `r names(read_csv("./fivethirtyeight_datasets/snp.csv"))`. By applying the similar cleaning step, this data set is turned into three variables, which are `r names(snp_df)`. The final dimension (rows, columns) of this data set is (`r dim(snp_df)`). 
* The original unemployment data set contains 13 variables, which are `r names(unemployment_df)`. Then the wide format is turned into the long format so that we have three variables, which are `r names(unemployment_df_tidy)`. The final dimension (rows, columns) of this data set is (`r dim(unemployment_df_tidy)`).
* The dimension (rows, columns) of the final data set is (`r dim(pols_snp_unemployment_df)`). The range of the year is (`r range(pull(pols_snp_unemployment_df, year))`). This final data set contains 11 variables, which are `r names(pols_snp_unemployment_df)`. "year" and "month" are key variables.